---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# presize

**Warning: R-package in development.**

The goal of presize is to provide functions for precision based sample size calculation. For a given sample size, the functions will return the precision (half the width of the confidence interval), and vice versa.

## Installation

You can install presize from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("a-lenz/presize")
```

## Overview
presize will provide functions for

* descriptive statistics
    + mean (`prec_mean`)
    + proportion (`prec_prop`)
    + rate (`prec_rate`)
  
* absolute and relative differences
    + mean difference
    + risk difference
    + odds ration
    + risk ratio
    + rate ratio
    + hazard ratio
  
* correlation measures
    + correlation coefficient
    + Cohens kappa
    + ICC
    + limit of agreement from Bland Altman plot
  
* diagnostic measures
    + sens
    + spez
    + positive LR
    + negative LR
    + AUC

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(presize)

# calculate sample size for a proportion of 0.2 with a precision of 0.2
prec_prop(p = .2, prec = .2, method = "wilson")
```
